---
title: "Quick `R` package"
author: "Laurent Gatto - http://github.com/lgatto" 
date: "7 July 2015 - CambR meeting"
output: ioslides_presentation
---


## Intro
- So easy: many benefits, little overhead
- Science: still loose code submitted/published
- Better `R` software engineering

### Why/what do you get
- Meta-data, versioning, dependencies handling
- Documentation, data, testing
- [Software|Data|Annotation/reports] packages
- **Project management**

## Advice
- Start early
- Document early
- Test it!
- Split personality: *developer* and *user*

## Example

```{r}
f <- function() message("Hello CambR")
f()
```

```{r, eval=FALSE}
package.skeleton("qp", list = "f") ## only once
devtools::create("qp") 
```

(RStudio has a package creation wizard.)

## A package is
- Meta-data
- code (`R`, but also `C`, `C++`, FORTRAN, ...)
- Documentation
- (data)
- ...

### Development cycle

```
R CMD build qp
R CMD check qp_0.1.tar.gz
R CMD INSTALL qp_0.1.tar.gz
```

## Vignettes

- Add an `Rmd` file in `qp/vignettes/`.
- Document that we need `knitr` and `rmarkdown::render` to produce the html file

```
Suggests: knitr, rmarkdown
VignetteBuilder: knitr
```

## Depending on other packages

In the `DESCRIPTION` file:

- Depends: loaded and attached - *user*
- Imports: loaded only - *developer*
- Suggests: needed for checking or optional

## Development cycle

Simplify build/check/install cycle with

```{r, eval=FALSE}
devtools::load_all("qp")
```


## Inline docs with `roxygen2`

```{r}
##' The description of the function comes here.
##'
##' @title A friendly message
##' @return \code{NULL}; used for its side effect.
##' @author Laurent Gatto
f <- function() message("Hello CambR")
```

```{r, eval=FALSE}
roxygen2::roxygenize('qp') 
roxygen2::roxygenize('qp', roclets='rd') 
```

## Unit testing
- `RUnit`
- `testthat`

```{r, echo=FALSE}
library("testthat")
```

```{r testthat}
## in qp/test/testthat/test_f.R
test_that("Test f function", {
              expect_null(f())
              expect_message(f(), "Hello CambR")
          })
```

```{r err, error=TRUE}
expect_message(f(), "Hello CambR!")
```

```
Suggests: testthat
```

## Coverage
```{r cov, eval=FALSE}
pc <- covr::package_coverage("qp")
shine(pc)
```

## Where
- CRAN - `install.packages`
- Bioconductor - `BiocInstaller::biocLite`
- github - `devtools::install_github`

### References

- [Writing `R` Extensions](http://cran.r-project.org/doc/manuals/r-devel/R-exts.html), R Core Team
- [`R` package](http://r-pkgs.had.co.nz/), Hadley Wickham

